---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# enrichr.db

The goal of enrichr.db is to provide the data from [Enrichr](http://amp.pharm.mssm.edu/Enrichr/) in
a format that makes it easy to access the gene sets programmatically from R.

The current Enrichr release included is from 2019-01-23.

## Installation

You can install the released version of enrichr.db from
[Github](https://github.com/clemenshug/enrichr.db) with:

```{r installation, eval = FALSE}
if (!requireNamespace("remotes", quietly = TRUE))
  install.packages("remotes")
remotes::install_url("https://github.com/clemenshug/enrichr.db/releases/download/v0.1/enrichr.db_0.1.tar.gz")
```

## Example

This is a basic example which shows you how to query the gene set database:

```{r query_drug_sets}
library(tidyverse)
library(enrichr.db)

# Finding all libraries that have something to do with drug treatments
drug_gene_sets <- enrichr_terms %>%
  filter(grepl("drug", library, ignore.case = TRUE))
```

This gives a data frame of gene set libraries. The gene sets are in the `data`
column of the data frame and are implemented as lists. Each list element is
a gene set and contains a vector of HGCN gene symbols.

```{r echo = FALSE}
knitr::kable(drug_gene_sets %>% mutate(data = paste0("<", map_int(data, length), " gene sets>")))
```

```{r print_drug_sets}
# Printing 10 genes of the first 5 gene sets in the first library we found (Drug_Perturbations_from_GEO_2014)
print(
  head(drug_gene_sets$data[[1]], n = 5) %>%
    map(head, n = 10)
)
```

## Funding

We gratefully acknowledge support by NIH Grant 1U54CA225088-01: Systems
Pharmacology of Therapeutic and Adverse Responses to Immune Checkpoint and Small
Molecule Drugs.
